---
title: "data 607 project 2"
author: "Yina Qiao"
date: "2023-03-04"
output:
  html_document:
    df_print: paged
---

## Introduction
The goal of this assignment is to practice in preparing different datasets for downstream analysis work

## DataSet 1 NYC MTA Subway Ridership from 2013

## load packages
```{r load packages, message=FALSE}
# Load required libraries
library(tidyverse)
```

## load data
```{r}
url <- 'https://raw.githubusercontent.com/yinaS1234/data-607/main/project%202/Annual20Subway20Ridership.csv'

dfMTA <- read.csv(file = url)

glimpse(dfMTA)
```
## Tidy Data
```{r}
# changing the column names
new_col_name <- c('Station', 2013, 2014, 2015, 2016, 2017, 2018, '2017 - 2018 Net Change', 
                   '2017 - 2018 % Change', '2018 Rank')
colnames(dfMTA) <- new_col_name

# finding the rows where the boroughs are entered
borough <- c('The Bronx', 'Brooklyn', 'Manhattan', 'Queens')

rowvalues <- c()

for(i in 1:length(borough)){
  rowvalues[i] <- rownames(dfMTA[which(dfMTA$'2013' == borough[i]),])
}
rowvalues
```

```{r}
#now that we now where the boroughs dataset begins and ends, we can capture the 
# data accordingly
dfBronx <- dfMTA[2:69,]
dfBronx['Borough'] <- borough[1]
  
dfBrooklyn <- dfMTA[71:227,]
dfBrooklyn['Borough'] <- borough[2]

dfManhattan <- dfMTA[229:349,]
dfManhattan['Borough'] <- borough[3]

dfQueens <- dfMTA[351:dim(dfMTA)[1],]
dfQueens['Borough'] <- borough[4]

# combined all sub datasets
dfMTA2 <- rbind(dfBronx, dfBrooklyn, dfManhattan, dfQueens)

# changed the columns from character to integer and removing commas
dfMTA2 <- dfMTA2 %>%
  mutate('2013' = as.integer(str_remove_all(dfMTA2$'2013', ',')),
         '2014' = as.integer(str_remove_all(dfMTA2$'2014', ',')),
         '2015' = as.integer(str_remove_all(dfMTA2$'2015', ',')),
         '2016' = as.integer(str_remove_all(dfMTA2$'2016', ',')),
         '2017' = as.integer(str_remove_all(dfMTA2$'2017', ',')),
         '2018' = as.integer(str_remove_all(dfMTA2$'2018', ',')),
         '2017 - 2018 Net Change' = as.integer(str_remove_all(dfMTA2$'2017 - 2018 Net Change', ',')),
         '2017 - 2018 % Change' = as.numeric(str_remove_all(dfMTA2$'2017 - 2018 % Change', '%')),
         '2018 Rank' = as.integer(dfMTA2$'2018 Rank')) %>%
  select(Borough, colnames(dfMTA2))
```

## Data Analysis

Let's look at the data by boroughs. 

```{r}

# subset of the data we want to look at
colnames2 <- c('Borough', 2013, 2014, 2015, 2016, 2017, 2018)
  
dfMTA3 <- dfMTA2 %>%
  select(colnames2)

aggMTA <- dfMTA3 %>%
  pivot_longer(!Borough, names_to = 'Year', values_to = 'Ridership') %>%
  group_by(Borough, Year) %>%
  summarize(Avg_Ridership = mean(Ridership, na.rm = TRUE))

ggplot(data = aggMTA) +
  geom_bar(mapping = aes(x = Year, y = Avg_Ridership, fill = Borough), stat = 'identity') +
  facet_grid(~ Borough) +
  theme(axis.text.x = element_text(angle = 70, hjust = 1)) +
  labs(title = 'Average Riderhip from 2013 - 2018 by Borough')

```

## Conclusion for Dataset 1
There is only minor change for ridership by boroughs from 2013 - 2018. The Queens borough has the most riders among all other boroughs.


## DataSet 2 MoviesOnStreamingPlatforms

I am curious to see which platform has better movies. 

##load data

```{r}
url <- 'https://raw.githubusercontent.com/yinaS1234/data-607/main/project%202/MoviesOnStreamingPlatforms.csv'

dfMovies <- read.csv(file = url)
glimpse(dfMovies)
```


## Tidy Data and data manipulation
```{r}
# Cleaning
dfMovies <- dfMovies[,-1] %>%
  rename(Rotten_Tomatoes = Rotten.Tomatoes, Prime_Video = Prime.Video, Disney = Disney.) %>%
  mutate(Rotten_Tomatoes = as.integer(str_remove(Rotten_Tomatoes, '%')))

# Transforming - we need to identify the platforms where the movies can be streamed.
# I created a subset for each platform and then combined them after
dfNetflix <- dfMovies %>%
  filter(Netflix == 1) %>%
  select(Title, IMDb, Rotten_Tomatoes)
dfNetflix['Platform'] <- 'Netflix'

dfHulu <- dfMovies %>%
  filter(Hulu == 1) %>%
  select(Title, IMDb, Rotten_Tomatoes)
dfHulu['Platform'] <- 'Hulu'

dfPrime_Video <- dfMovies %>%
  filter(Prime_Video == 1) %>%
  select(Title, IMDb, Rotten_Tomatoes)
dfPrime_Video['Platform'] <- 'Prime_Video'

dfDisney <- dfMovies %>%
  filter(Disney == 1) %>%
  select(Title, IMDb, Rotten_Tomatoes)
dfDisney['Platform'] <- 'Disney'

dfMovies2 <- rbind(dfNetflix, dfHulu, dfPrime_Video, dfDisney)


```

## Data Analysis

```{r}
ggplot(data = dfMovies2, aes(x = Platform, y = IMDb, fill = Platform)) +
  geom_boxplot() + 
  labs(title = 'IMDb Rating by Platform') 


ggplot(data = dfMovies2, aes(x = Platform, y = Rotten_Tomatoes, fill = Platform)) +
  geom_boxplot() + 
  labs(title = 'Rotten Tomoatoes Rating by Platform') 
```

## Conclusion on Dataset 2

It seems like Rotton tomatoes rating has many NA and not enough data to support finding, where IMDb rating has less NA and therefore is a better indicator. Recommend to compare platforms based on IMDb rating, Disney has highest IMDb, while Prime-Video has lowest IMDb. 


## DataSet3 School Diversity

I am curious to see the difference racial average per school.  

##load data
```{r}
url <- 'https://raw.githubusercontent.com/yinaS1234/data-607/main/project%202/School_Diversity.csv'

dfSchool <- read.csv(file = url)
str(dfSchool)
```

## Data Cleaning and Manipulation

```{r}
dfSchool2 <- dfSchool %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  filter(Total > 100) %>%
  mutate(Asian = Asian / 100,
         Black = Black / 100,
         Hispanic = Hispanic / 100,
         White = White / 100,
         Multi = Multi / 100,
         ) %>%
  group_by(ST) %>%
  summarize(Asian = mean(Asian),
            Black = mean(Black),
            Hispanic = mean(Hispanic),
            White = mean(White),
            Multi = mean(Multi)) %>%
  pivot_longer(!ST, names_to = 'Ethnicity', values_to = 'Ratio')

```
```

## Data Analysis

The above data shows the ratio average across school.  Let's plot it for better analysis. 

```{r}
ggplot(data = dfSchool2, aes(x = Ethnicity, y = Ratio, fill = Ethnicity)) +
  geom_boxplot() +
  labs(title = 'Ethnicity Ratio Distribution in Schools')
```


## Conclusion Dataset 3 
From our analysis, we see that the overall the spread of White student ratio is the highest by a wide margin, where the other race ratios are more similars.  





