---
title: "data 607 lab3"
author: "Yina Qiao"
date: "2023-02-13"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## 1
```{r}
df.college.majors = read.csv( url("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"))

vec.majors = df.college.majors$Major[grep("DATA|STATISTICS", df.college.majors$Major)] 
print(vec.majors)
```

## 2
```{r}
vec.text = c('[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"')

vec.text.char = gsub("(\\n\\[\\d+\\])|(^\\[\\d+\\])", "", vec.text)
vec.text.char = strsplit(vec.text.char, '\\"')
vec.text.char = unlist(vec.text.char)
vec.text.char = vec.text.char[grep("[a-z]", vec.text.char)]
print(vec.text.char)
```

## 3

1> (.)\1\1. This will detect any three consecutive charactes in string format that has the same character.


```{r}
str_detect("AAA", "(.)\1\1")
```

```{r}
str_detect("A\1\1", "(.)\1\1")
```

2> "(.)(.)\\2\\1". This matches any 4 consecutive characters in a string where
the last 2 are the same as the first 2 characters in reverse order.
```{r}
str_detect("eppe","(.)(.)\\2\\1")
```


3> (..)\1
Regular expression not represented in string format that has two characters repeated twice in the same order
```{r}
str_detect('nana', '(..)\1')
```

4> "(.).\\1.\\1"
This will match any five consecutive characters where character 1, 3, and 5 are the same
```{r}
str_detect('abana',"(.).\\1.\\1")
```


5> "(.)(.)(.).*\\3\\2\\1"
match any string that contains at least six characters with the last 3 characters as the same as the first 3 characters in reverse order

```{r}
str_detect("123newyorkcity;321","(.)(.)(.).*\\3\\2\\1")
```



## 4
Construct regular expressions to match words that:
1>Start and end with the same character.
```{r}
#"^(.).*\\1$"
```

2>Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
```{r}
# "(..).*\\1"
```

3>Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
```{r}
# "(.).*\\1.*\\1"
```


