---
title: "Data 607 lab 5"
author: "Yina Qiao"
date: "2023-02-26"
output: html_document
---


## Introduction
This assignment will import uncleaned data from a csv. file. My task is to tidy and transform data as described below.
(1) Read the information from a .CSV file into R, and use tidyr and dplyr as needed to tidy and transform the data.
(2) Perform analysis to compare the arrival delays for the two airlines


## Install and load pakages
```{r}
library(tidyverse)

```

## Load dataset
```{r}
url = r"(https://raw.githubusercontent.com/yinaS1234/data-607/main/data%20607%20lab%205/lab5data.csv)"
suppressMessages(
df <- read_csv(url, skip_empty_rows = TRUE, show_col_types = FALSE)[-3,] 
)
names(df)[1:2] <- c("Airline", "Status")
df
```

## Clean and tidy the data
*fixing NA
transform data, rename*

```{r}
for(x in seq(from=2, to=nrow(df), by=2))
{
  df[x, 1] = df[x-1, 1]
}
```

```{r}
library(tidyr)

df_unpivot = df %>%
  gather(key="City", value="Count", c("Los Angeles","Phoenix", "San Diego", "San Francisco", "Seattle"))
df_unpivot
```

```{r}
library(dplyr)

df_unpivot_2 = df_unpivot %>%
  spread(key="Status", value="Count")

df_unpivot_2 = df_unpivot_2 %>%
  rename(on_time = `on time`)

```

## Data Analysis

### On Time Percentage By City by airline
```{r}
df_unpivot_2 = df_unpivot_2 %>%
  mutate(otp = on_time/(on_time + delayed)) %>%
  arrange(desc(otp))

df_unpivot_2
```

### Overall On Time Percentage By Airline.
* we see that AM West Outperform ALASKA overall.  However,could the result skewed due to more flights in certain city? 
```{r}
df_unpivot_2 %>%
  select(Airline, delayed, on_time) %>%
  group_by(Airline) %>%
  summarise(delayed = sum(delayed), on_time = sum(on_time), otp = sum(on_time) / (sum(on_time) + sum(delayed))) %>%
  arrange(desc(otp))
```

* it seems like AM WEST has more flights in Phoenix, so lets compute the OTP without phoenix.*
```{r}
df_unpivot_2 %>%
  filter(City != "Phoenix") %>%
  select(Airline, delayed, on_time) %>%
  group_by(Airline) %>%
  summarise(delay = sum(delayed), on_time = sum(on_time), otp = sum(on_time) / (sum(on_time) + sum(delayed))) %>%
  arrange(desc(otp))
```

## Conclusion

Based on the data provided,ALASKA consistently outperforms AM WEST. For flying to Phoenix specifically, I would recommend AM WEST as the on time performance is on par with ALASKA and have more flights available.



